# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
dat=read.csv('activity.csv');
dat=transform(dat,date=as.Date(date)) #transform the "date" column to the R class "Date"
```

## What is mean total number of steps taken per day?
```{r}
#aggregate is a handy function, like in SQL one can use "sum(xxx) GROUP BY yyy"
daysum=aggregate(dat$steps,by=list(dat$date),sum)
names(daysum)=c("day","steps")
hist(daysum[,2],20,xlab = "total number of steps taken",main="Histogram of total number of steps taken each day")
```
### The mean and median total number of steps taken per day
```{r}
print(paste("mean:",mean(daysum$steps, na.rm = TRUE)))
print(paste("median:",median(daysum$steps, na.rm = TRUE )))
```


## What is the average daily activity pattern?
```{r}
avgstep=aggregate(dat$steps,by=list(dat$interval),mean,na.rm=TRUE)
names(avgstep)=c('5 minute interval','average_steps')
plot(avgstep,type='l')
```
### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
avgstep[which.max(avgstep[,2]),1]
```


## Imputing missing values
- the total number of missing values in the dataset:
```{r}
sum(is.na(dat$steps))
```
-  Fill in missing values using the mean for the interval
```{r}
#merge is also a handy function, just like SQL join
dat2=merge(dat,avgstep,by.x="interval",by.y=names(avgstep)[1])
dat2$steps[is.na(dat2$steps)]<-dat2$average_steps[is.na(dat2$steps)]

```
The dataset after filling in the missing values:
```{r}
head(dat2)
```
### Histogram and mean and median after filling the missing values:
```{r}
daysum=aggregate(dat2$steps,by=list(dat2$date),sum)
names(daysum)=c("day","steps")
hist(daysum[,2],20,xlab = "total number of steps taken",main="Histogram of total number of steps taken each day")
print(paste("mean:",mean(daysum$steps, na.rm = TRUE)))
print(paste("median:",median(daysum$steps, na.rm = TRUE )))
```

The mean is equal to the origin one

The median becomes the same as mean

The histogram's shape is nearly unchanged.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
Sys.setlocale("LC_TIME", "English")

#creating a factor variable
dat2$weekdays=factor(format(dat2$date, "%A"))
levels(dat2$weekdays)

# replace the levels
levels(dat2$weekdays) <- list(weekday = c("Monday", "Tuesday", "Wednesday", 
    "Thursday", "Friday"), weekend = c("Saturday", "Sunday"))
levels(dat2$weekdays)
````

### Making plots
Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
library(lattice)
#Note: according to the way that I generated the column "average_steps"(using "merge" function), it is exactly what I want to plot
xyplot(average_steps~interval | weekdays,data=dat2,layout=c(1,2),type='l')
```